---
title: "PracticalMLProject"
author: "Apoorv Kulshreshtha"
date: "Saturday, July 26, 2014"
output: html_document
---
  In my first attempt at the problem, I found it a little difficult to comprehend and relate the "pml-training.csv" data with the problem statement.I then visited the website which was the source of the data given in the problem: http://groupware.les.inf.puc-rio.br/har .The information provided on this website under the "Weight Lifting Exercises Dataset" heading, helped me understand the objective of the question much better.The first inference that I made after going through the data was that many columns had zero/NA values for most of the observations. Therefore, I decided to drop those columns and check if the accuracy of the model is hampered by doing so.If the accuracy would have been significantly reduced, I would have reverted back to the complete data.Once the data had been cleaned, I used random forest to predict the classe variable. To prevent overfitting of the model on the training data, train control method "cv" was used which does the implicit cross-validation.The reference which was used to understand the train control method is : http://caret.r-forge.r-project.org/training.html .The accuracy achieved by the model while predicting the test data was approximately 0.99, making the out of sampe error in cross validation close to 0.01. The accuracy achieved by the model proved that the columns with zero/NA value were indeed insignificant, and dropping them from the analysis had little impact on the model.Following is the summary of the model used for prediction purpose:

```{r, echo=FALSE}
library(caret)
InitialTrainingData <- read.csv("F:/Acads/PracticalMachineLearning/pml-training.csv")
FinalTrainingData <- subset(InitialTrainingData,select = c(roll_belt,pitch_belt,yaw_belt,total_accel_belt,gyros_belt_x,gyros_belt_y,gyros_belt_z,accel_belt_x,accel_belt_y,accel_belt_z,magnet_belt_x,magnet_belt_y,magnet_belt_z,roll_arm,pitch_arm,yaw_arm,total_accel_arm,gyros_arm_x,gyros_arm_y,gyros_arm_z,accel_arm_x,accel_arm_y,accel_arm_z,magnet_arm_x,magnet_arm_y,magnet_arm_z,roll_dumbbell,pitch_dumbbell,yaw_dumbbell,total_accel_dumbbell,gyros_dumbbell_x,gyros_dumbbell_y,gyros_dumbbell_z,accel_dumbbell_x,accel_dumbbell_y,accel_dumbbell_z,magnet_dumbbell_x,magnet_dumbbell_y,magnet_dumbbell_z,roll_forearm,pitch_forearm,yaw_forearm,total_accel_forearm,gyros_forearm_x,gyros_forearm_y,gyros_forearm_z,accel_forearm_x,accel_forearm_y,accel_forearm_z,magnet_forearm_x,magnet_forearm_y,magnet_forearm_z,classe))
myControl <- trainControl(method='cv')
PredictionModel <- train(classe~.,data=FinalTrainingData,method="rf",trcontrol= myControl)
PredictionModel
InitialTestData <- read.csv("F:/Acads/PracticalMachineLearning/pml-testing.csv")
FinalTestData <- subset(InitialTestData,select = c(roll_belt,pitch_belt,yaw_belt,total_accel_belt,gyros_belt_x,gyros_belt_y,gyros_belt_z,accel_belt_x,accel_belt_y,accel_belt_z,magnet_belt_x,magnet_belt_y,magnet_belt_z,roll_arm,pitch_arm,yaw_arm,total_accel_arm,gyros_arm_x,gyros_arm_y,gyros_arm_z,accel_arm_x,accel_arm_y,accel_arm_z,magnet_arm_x,magnet_arm_y,magnet_arm_z,roll_dumbbell,pitch_dumbbell,yaw_dumbbell,total_accel_dumbbell,gyros_dumbbell_x,gyros_dumbbell_y,gyros_dumbbell_z,accel_dumbbell_x,accel_dumbbell_y,accel_dumbbell_z,magnet_dumbbell_x,magnet_dumbbell_y,magnet_dumbbell_z,roll_forearm,pitch_forearm,yaw_forearm,total_accel_forearm,gyros_forearm_x,gyros_forearm_y,gyros_forearm_z,accel_forearm_x,accel_forearm_y,accel_forearm_z,magnet_forearm_x,magnet_forearm_y,magnet_forearm_z))
FinalPrediction <- predict(PredictionModel,FinalTestData)
```

The above model when used for predicting the test data, gave the following result:
```{r, echo=FALSE}
FinalPrediction
```
